Formal Synthesis of Digital Controllers for Observable Continuous Plants with Transient Performance Specifications

### 2018-04-11 CAV2018/paper.tex
corrections based on cav2018 reviews

While research on digital control is well
developed~\cite{astrom1997computer}, automated and sound control synthesis
is challenging, particularly when the synthesis objective goes beyond
classical stability.  Recent work~\cite{abate2017automated} described a
method for synthesizing safe digital stabilizing controllers for fully
observable LTI plants.  This work differs from the vast majority of existing
approaches in that it explores the continuous space of the system to
synthesize correct-by-construction minimal digital controllers.  In contrast
to LTL models, which require thousands of states and elaborate algorithms to
implement~\cite{reissig2017feedback}, these PID-like controllers are
represented by a few parameters and only require few lines of code, thus
reducing the implementation cost and execution time while remaining
resilient to the effects of digitalization. 
However,~\cite{abate2017automated}~is restricted to fully observed systems
(i.e., where all the states are known) modeled in discrete time, and
furthermore explores only a subset of the properties that engineers are
interested in.  By contrast, in most real implementations, the internal
state of a plant cannot be directly measured but has to be inferred from its
effect on the output.  A plant model that minimizes the difference between
its output and the real output of the plant is called an \emph{observer} and
it is a well known paradigm in the control
literature~\cite{astrom1997computer}.

@@

We use a CounterExample-Guided Inductive Synthesis (CEGIS)~\cite{jha-icse10,
DBLP:conf/pldi/Solar-LezamaRBE05} approach, which we enhance with
abstraction refinement and optimization resulting in the novel CEGIS-OR
(CEGIS with Optimization and Refinement).  This design uses an iterative
process where each iteration performs inductive generalization based on
counterexamples provided by a verification oracle.  Essentially, the
inductive generalization uses information about a limited number of inputs
to make claims in the form of candidate solutions about all the possible
inputs.  By~using abstraction refinement, CEGIS-OR allows us to handle
complex systems, which are initially represented by coarse abstractions that
only get refined on demand.  Consequently, solutions that do not require a
high level of precision are obtained fast.  Since the use of coarse
abstractions may result in low-quality candidate solutions (i.e., candidates
that fail the specification on a very large number of inputs), we use
optimization techniques after each inductive generalization phase to improve
the produced candidates.  In particular, our optimization procedure reduces
the size of the reach space of the closed loop system, while maintaining
other dynamical properties of interest (safety, stability, and response
time).

@@

We are interested in control systems with an output $\vec{y}$ that is a
linear combination of its states and inputs\footnote{Notice that since we
can only measure $\vec{y}$, the state-space $\vec{x}$ is not fully
observable.}.
%
%, which may restrict the observability of the state-space from the output space.
\begin{equation}
\label{eq:ode2}  
\vec{y}(t)=\mat{C}\vec{x}(t)+\mat{D}\vec{u}(t),
\vec{y} \in \mathbb{R}^o,
\mat{C} \in \mathbb{R}^{o \times p},
\mat{D} \in \mathbb{R}^{o \times q}. 
\end{equation}
%
Equations~\eqref{eq:ode1} and \eqref{eq:ode2} are soundly discretized in
time~\cite{middleton1990digital,van1978computing} at constant intervals,
whose duration we refer to as the sample time $T_s$, into the difference
equation
%
\begin{align}
\label{eq:plant}
\left\{
\begin{array}{rl}
\vec{x}_{k+1} =& \mat{A}_d \vec{x}_k+ \mat{B}_d \vec{u}_k\\
\vec{y}_{k} =& \mat{C} \vec{x}_k + \mat{D} \vec{u}_k, 
\end{array}
\right.
\text{ where }
\left\{
\begin{array}{l}
\mat{A}_d=e^{\mat{A}T_s}\\
\mat{B}_d = \int_{t = 0}^{T_s} e^{\mat{A} t} dt\ \mat{B}
\end{array}
\right.,
\end{align} 
%
%where $\mat{A}_d=e^\mat{A}$, $\mat{B}_d = \int_{t = 0}^{T_s} e^{\mat{A} t} dt\ \mat{B}$,
%
where $k \in \mathbb N$ is a discrete counter and $\vec{x}_{0}=\vec{x}(0)$
is the initial state.  In this work we apply the standard assumption that
$\mat{D}=\mat{0}$~\cite{Astrom08}.
%
%Also, since $\vec{y}$ can be directly derived from $\vec{x}$, we will omit discussing it in the following unless strictly necessary.

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Semantics of LTI models}\label{sec:model_semantics}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
%
The traces of model \eqref{eq:plant}, starting from an initial set
$X_0\subseteq \mathbb{R}^p$, with inputs taken from the set $U \subseteq
\mathbb{R}^q$, are sequences
%
$ \vec{x}_0 \xrightarrow{\vec{u}_0} \vec{x}_1 \xrightarrow{\vec{u}_1} \vec{x}_2 \xrightarrow{\vec{u}_2} \ldots $
%
where
%
$ \vec{x}_0 \in X_0$ and $\forall k\geq 0, \vec{u}_k \in U \text{ and } \vec{x}_{k+1} = \tau(\vec{x}_k,\vec{u}_k) $, 
where 
%
\begin{equation}\label{equ:reachtraj}
\tau(\vec{x}_k,\vec{u}_k) = %\big\{ 
\mat{A}\vec{x}_k + \mat{B}\vec{u}_k \;.
 %\mid \mat{G}\vec{x}_k \leq \vec{h}\big\}\;. 
\end{equation}
%\note{do we actually need to mention the safety spec here given that it is formally introduced only in Sec~\ref{sec:cof_safety}?}
%
%Here $\mat{G}$ and $\vec{h}$ define a set of inequalities that describe a safe set, within the state-space $\mathbb{R}^p$.
%
We generalize this notation to convex sets of initial conditions and inputs
($X_0$ and $U$, respectively), and write $\tau(X_0,U)$ to denote the set of
states $\{\vec{x} \mid \vec{x} = \tau(\vec{x}_0,\vec{u}),\, \vec{x}_0 \in
X_0,\, \vec{u} \in U \}$ reached from $X_0$ by $\tau$ in one step.
%
We furthermore define the $n$-reach set $\tau^n(X_0,U)$ as the set of states
reached from $X_0$ via $\tau$ in $n$ steps.
%
\begin{align}\label{equ:reachset}
\tau^n(X_0,U) = \{\vec{x}_n \mid & 
\forall k\in [0,n-1] \, \vec{x}_{k+1}=\tau(\vec{x}_{k},\vec{u}_{k}), \,
\vec{u}_{k} \in U, \,   
\vec{x}_0 \in X_0
\} \;. 
\end{align}
%
Since the transformations $\mat{A}$ and $\mat{B}$ are linear and linear
transformations preserve convexity~\cite{boyd2004convex}, the sets $X_n =
\tau^n(X_0,U)$ are also convex.  We define the \emph{$n$-reach tube} as
%
\begin{equation}\label{equ:reachtube}
\hat{X}_n=\hat{\tau}^n(X_0,U)=\bigcup_{k\in[0,n]} \tau^k(X_0,U)
\end{equation}
%
as the union of $k$-reach sets over $n$ iterations. 
%
Finally, the set $\hat{X} =\bigcup_{n\geq 0} \tau^n(X_0,U)$, which we refer
to as the reach space of the system, generalizes the previous notions over
an unbounded time horizon (transitive closure).

@@

The final case we analyze is that of hybrid-time models, where the plant
dynamics evolve in continuous time while the feedback dynamics evolve in
discrete time.  A~new input value is produced only at every sample time but
implemented over the continuous-time plant dynamics.
%
\begin{theorem}
The expression
%
 \begin{align}
 \vec{x}_{T} &= (\mat{A}_{T-kT_s}-\mat{B}_{T-kT_s}\mat{K}) (\mat{A}_d-\mat{B}_d\mat{K})^k\vec{x}_0, 
 \label{eq:cyber_feedback}
 \end{align}
%
with $\mat{A}_T$ as in \eqref{eq:continuous_tube_dyn2}, $\mat{B}_T$ as in
\eqref{eq:continuous_tube_param}, and where $\mat{A}_d=\mat{A}_{T_s},
\mat{B}_d=\mat{B}_{T_s}$ describe the discretized plant dynamics with
sample time $T_s$, is the state of the model
%
\begin{align*}
 \dot{\vec{x}}(t) = \mat{A}\vec{x}(t)+\mat{B}\vec{u}(t), \quad 
 \vec{u}(t)=-\mat{K}\vec{x}_k,  \,
 kT_s \leq t < (k+1)T_s \,. 
\end{align*}
%
For the proof see Appendix~\ref{sec:proofs}.
%
\end{theorem}
%\begin{proof}
%See Appendix~\ref{sec:proofs}
%\end{proof}

@@

\begin{theorem}
%
The error introduced by a delay $t_d$ in the control output $\vec{u}_k$ of a
discrete feedback controller for a continuous LTI system can be modeled as a
non-deterministic disturbance signal ($\nu_t$ in
Figure~\ref{fig:observersystem}) upper bounded by
%
\begin{align}
\norm{\vec{y}_{k}-{\vec{y}_{t_d}}_{k}}_1 \leq q_t, \text{ where } q_t=&(\norm{\mat{C}(\mat{B}_{T_s}-\mat{B}_{t_s})}_1+\norm{\mat{C}\mat{A}_{t_s}\mat{B}_{t_d}}_1)\norm{U}_1
\end{align}
where $\vec{y}_{t_d}$ is the output of the system with delay $t_d$, $t_s=T_s-t_d$,
and $\norm{\vec{x}}_1=\sum_{i=0}^{size(\vec{x})} |\vec{x}_i|$ is the $l_1$ norm, 
which we extend to sets as $\norm{X}_1 = \sup \{\norm{\vec{x}}_1, \vec{x} \in X\}$.  
For the proof see Appendix~\ref{sec:proofs}
\end{theorem}
%
%\begin{proof}
%See Appendix~\ref{sec:proofs}
%\end{proof}
This result allows to over-approximate the effects of the delay by introducing a disturbance signal $\nu_t=[-q_t, q_t]$ at the output.
 %and use the dynamics of the system without delay.  
 This noise will be small when the delay is much smaller than the sampling time ($t_s\approx T_s$ and $\mat{B}_{t_d} \approx 0$).

>>>>

While research on digital control is well
developed~\cite{astrom1997computer}, automated and sound control synthesis
is challenging, particularly when the synthesis objective goes beyond
classical stability.  Recent work~\cite{abate2017automated} described a
method for synthesizing safe digital stabilizing controllers for fully
observable LTI plants.  This work differs from the vast majority of existing
approaches in that it explores the continuous space of the system to
synthesize correct-by-construction minimal digital controllers.  In contrast
to LTL models, which require thousands of states and elaborate algorithms to
implement~\cite{reissig2017feedback}, these PID-like controllers are
represented by a few parameters and only require few lines of code, thus
reducing the implementation cost and execution time while remaining
resilient to the effects of digitalization. 
However,~\cite{abate2017automated}~is restricted to fully observed systems
(i.e., where all the states are known) modeled in discrete time, and
furthermore explores only a subset of the properties that engineers are
interested in.  By contrast, in most real implementations, the internal
state of a plant cannot be directly measured but has to be inferred from its
effect on the output. A system that converges to the internal state of the plant
from measurements of its input and output is called an \emph{observer}, and
it is a well known paradigm in the control literature~\cite{astrom1997computer}.

@@

We use a CounterExample-Guided Inductive Synthesis (CEGIS)~\cite{jha-icse10,
DBLP:conf/asplos/Solar-LezamaTBSS06} approach, which we enhance with
abstraction refinement and optimization resulting in the novel CEGIS-OR
(CEGIS with Optimization and Refinement).  This design uses an iterative
process where each iteration performs inductive generalization based on
counterexamples provided by a verification oracle.  Essentially, the
inductive generalization uses information about a limited number of inputs
to make claims in the form of candidate solutions about all the possible
inputs.  By~using abstraction refinement, CEGIS-OR allows us to handle
complex systems, which are initially represented by coarse abstractions that
only get refined on demand.  Consequently, solutions that do not require a
high level of precision are obtained fast.  Since the use of coarse
abstractions may result in low-quality candidate solutions (i.e., candidates
that fail the specification on a very large number of inputs), we use
optimization techniques after each inductive generalization phase to improve
the produced candidates.  In particular, our optimization procedure reduces
the size of the reach space of the closed loop system, while maintaining
other dynamical properties of interest (safety, stability, and response
time).

@@

We are interested in control systems with an output $\vec{y}$ that is a
linear combination of its states and inputs\footnote{Notice that since we
can only measure $\vec{y}$, the state $\vec{x}$ cannot be measured directly.}.
%
%, which may restrict the observability of the state-space from the output space.
\begin{equation}
\label{eq:ode2}  
\vec{y}(t)=\mat{C}\vec{x}(t)+\mat{D}\vec{u}(t),
\vec{y} \in \mathbb{R}^o,
\mat{C} \in \mathbb{R}^{o \times p},
\mat{D} \in \mathbb{R}^{o \times q}. 
\end{equation}
%
Equations~\eqref{eq:ode1} and \eqref{eq:ode2} are soundly discretized in
time~\cite{middleton1990digital,van1978computing} at constant intervals,
whose duration we refer to as the sample time $T_s$, into the difference
equation
%
\begin{align}
\label{eq:plant}
\left\{
\begin{array}{rl}
\vec{x}_{k+1} =& \mat{A}_d \vec{x}_k+ \mat{B}_d \vec{u}_k\\
\vec{y}_{k} =& \mat{C} \vec{x}_k + \mat{D} \vec{u}_k, 
\end{array}
\right.
\text{ where }
\left\{
\begin{array}{l}
\mat{A}_d=e^{\mat{A}T_s}\\
\mat{B}_d = \int_{t = 0}^{T_s} e^{\mat{A} t} dt\ \mat{B}
\end{array}
\right.,
\end{align} 
%
%where $\mat{A}_d=e^\mat{A}$, $\mat{B}_d = \int_{t = 0}^{T_s} e^{\mat{A} t} dt\ \mat{B}$,
%
where $k \in \mathbb N$ is a discrete counter and $\vec{x}_{0}=\vec{x}(0)$
is the initial state.  In this work we apply the standard assumption that
$\mat{D}=\mat{0}$~\cite{Astrom08}.
%
%Also, since $\vec{y}$ can be directly derived from $\vec{x}$, we will omit discussing it in the following unless strictly necessary.

%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
\subsection{Semantics of LTI models}\label{sec:model_semantics}
%+++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
%
The traces of model \eqref{eq:plant}, starting from an initial set
$X_0\subseteq \mathbb{R}^p$, with inputs taken from the set $U \subseteq
\mathbb{R}^q$, are sequences
%
$ \vec{x}_0 \xrightarrow{\vec{u}_0} \vec{x}_1 \xrightarrow{\vec{u}_1} \vec{x}_2 \xrightarrow{\vec{u}_2} \ldots $
%
where
%
$ \vec{x}_0 \in X_0$ and $\forall k\geq 0, \vec{u}_k \in U \text{ and } \vec{x}_{k+1} = \tau(\vec{x}_k,\vec{u}_k) $, 
where 
%
\begin{equation}\label{equ:reachtraj}
\tau(\vec{x}_k,\vec{u}_k) = %\big\{ 
\mat{A}_d\vec{x}_k + \mat{B}_d\vec{u}_k \;.
 %\mid \mat{G}\vec{x}_k \leq \vec{h}\big\}\;. 
\end{equation}
%\note{do we actually need to mention the safety spec here given that it is formally introduced only in Sec~\ref{sec:cof_safety}?}
%
%Here $\mat{G}$ and $\vec{h}$ define a set of inequalities that describe a safe set, within the state-space $\mathbb{R}^p$.
%
We generalize this notation to convex sets of initial conditions and inputs
($X_0$ and $U$, respectively), and write $\tau(X_0,U)$ to denote the set of
states $\{\vec{x} \mid \vec{x} = \tau(\vec{x}_0,\vec{u}),\, \vec{x}_0 \in
X_0,\, \vec{u} \in U \}$ reached from $X_0$ by $\tau$ in one step.
%
We furthermore define the $n$-reach set $\tau^n(X_0,U)$ as the set of states
reached from $X_0$ via $\tau$ in $n$ steps.
%
\begin{align}\label{equ:reachset}
\tau^n(X_0,U) = \{\vec{x}_n \mid & 
\forall k\in [0,n-1] \, \vec{x}_{k+1}=\tau(\vec{x}_{k},\vec{u}_{k}), \,
\vec{u}_{k} \in U, \,   
\vec{x}_0 \in X_0
\} \;. 
\end{align}
%
Since the transformations $\mat{A}$ and $\mat{B}$ are linear and linear
transformations preserve convexity~\cite{boyd2004convex}, the sets $X_n =
\tau^n(X_0,U)$ are also convex.  We define the \emph{$n$-reach tube} as
%
\begin{equation}\label{equ:reachtube}
\hat{X}_n=\hat{\tau}^n(X_0,U)=\bigcup_{k\in[0,n]} \tau^k(X_0,U)
\end{equation}
%
as the union of $k$-reach sets over $n$ iterations. 
%
Finally, the set $\hat{X} =\bigcup_{n\geq 0} \tau^n(X_0,U)$, which we refer
to as the reach space of the system, generalizes the previous notions over
an unbounded time horizon (transitive closure).

@@

The final case we analyze is that of hybrid-time models, where the plant
dynamics evolve in continuous time while the feedback dynamics evolve in
discrete time.  A~new input value is produced only at every sample time but
implemented over the continuous-time plant dynamics.
%
\begin{theorem}
The expression
%
 \begin{align}
 \vec{x}_{T} &= (\mat{A}_{T-kT_s}-\mat{B}_{T-kT_s}\mat{K}) (\mat{A}_d-\mat{B}_d\mat{K})^k\vec{x}_0, 
 \label{eq:cyber_feedback}
 \end{align}
%
with $\mat{A}_T$ as in \eqref{eq:continuous_tube_dyn2}, $\mat{B}_T$ as in
\eqref{eq:continuous_tube_param}, $T \in [kT_s, (k+1)T_s]$, and where $\mat{A}_d=\mat{A}_{T_s},
\mat{B}_d=\mat{B}_{T_s}$ describe the discretized plant dynamics with
sample time $T_s$, is the state of the model
%
\begin{align*}
 \dot{\vec{x}}(t) = \mat{A}\vec{x}(t)+\mat{B}\vec{u}(t), \quad 
 \vec{u}(t)=-\mat{K}\vec{x}_k,  \,
 kT_s \leq t < (k+1)T_s \,. 
\end{align*}
%
For the proof see Appendix~\ref{sec:proofs}.
%
\end{theorem}
%\begin{proof}
%See Appendix~\ref{sec:proofs}
%\end{proof}

@@

\begin{theorem}
%
The error introduced by a delay $t_d$ in the control input $\vec{u}_k$ updated
by a discrete feedback controller for a continuous LTI system can be modeled as
a non-deterministic disturbance signal ($\nu_t$ in
Figure~\ref{fig:observersystem}) upper bounded by
%
\begin{align}
\norm{\vec{y}_{k}-{\vec{y}_{t_d}}_{k}}_1 \leq q_t, \text{ where } q_t=&(\norm{\mat{C}(\mat{B}_{T_s}-\mat{B}_{t_s})}_1+\norm{\mat{C}\mat{A}_{t_s}\mat{B}_{t_d}}_1)\norm{U}_1
\end{align}
where $\vec{y}_{t_d}$ is the output of the system with delay $t_d$, $t_s=T_s-t_d$,
and $\norm{\vec{x}}_1=\sum_{i=0}^{size(\vec{x})} |\vec{x}_i|$ is the $l_1$ norm, 
which we extend to sets as $\norm{X}_1 = \sup \{\norm{\vec{x}}_1, \vec{x} \in X\}$.  
For the proof see Appendix~\ref{sec:proofs}
\end{theorem}
%
%\begin{proof}
%See Appendix~\ref{sec:proofs}
%\end{proof}
This result allows to over-approximate the effects of the delay by introducing a disturbance signal $\nu_t=[-q_t, q_t]$ at the output.
 %and use the dynamics of the system without delay.  
 This noise will be small when the delay is much smaller than the sampling time ($t_s\approx T_s$ and $\mat{B}_{t_d} \approx 0$).
